---
title: "Multi-Indicator Small Area Estimation"
subtitle: "`r params$country`"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  bookdown::pdf_book:
    toc: true
    number_sections: false
    keep_tex: true
header-includes:
  - '\usepackage{graphicx}'
  - '\usepackage{float}'
  - '\usepackage{booktabs}'
  - '\usepackage{longtable}'
  - '\usepackage{array}'
  - '\usepackage{multirow}'
  - '\usepackage{wrapfig}'
  - '\usepackage{colortbl}'
  - '\usepackage{pdflscape}'
  - '\usepackage{threeparttable}'
  - '\usepackage{threeparttablex}'
  - '\usepackage[normalem]{ulem}'
  - '\usepackage{makecell}'
  - '\usepackage{xcolor}'
  - '\setcounter{tocdepth}{2}'
  - '\usepackage{caption}'
  - '\captionsetup[table]{position=bottom}'
params:
  country: "Nigeria"
  source_path: null
  middle_path: "Gates-results/ReportPlots"
  yr1: null
  yr2: null
---
\clearpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.align = "center", comment = NA, cache = FALSE
)
library(knitr)

```

```{r}
# devtools::install_github("wu-thomas/sae4health")
library(geodata)
library(sf)
library(SUMMER)
library(survey)
library(rdhs)
library(ggplot2)
library(patchwork)
library(kableExtra)
library(dplyr)
library(tidyr)
library(terra)
library(raster)
library(surveyPrev)
library(qs)
library(knitr)
library(sae4health)
library(rmapshaper)
```


```{r}
source_path <- if (is.null(params$source_path)) dirname(here::here()) else params$source_path
country=params$country
middle_path=params$middle_path
git_path<- file.path(source_path,"SAE4LMIC")
infolist <- read.csv(file.path(git_path, "info", "infolist.csv"))
surveys <- read.csv(file.path(git_path,  "info", "surveyslist.csv"))
shapeused<- read.csv(file.path(git_path,  "info", "shapefileList.csv"))

```

```{r}

# old/new survey years: use params$yr1 / params$yr2 if supplied, else min/max
yrs_all <- sort(unique(shapeused[shapeused$country==country,]$year))
yr1 <- if (!is.null(params$yr1)) params$yr1 else min(yrs_all)
yr2 <- if (!is.null(params$yr2)) params$yr2 else max(yrs_all)


c1=shapeused[shapeused$country==country& shapeused$year==yr1,]$n_geo_cluster
c2=shapeused[shapeused$country==country& shapeused$year==yr2,]$n_geo_cluster



ad11=shapeused[shapeused$country==country& shapeused$year==yr1,]$n_admin1
ad12=shapeused[shapeused$country==country& shapeused$year==yr1,]$n_admin2
indicatorlist=infolist$ID



```


```{r}
get_image_files <- function(image_dir,
                            pattern,
                            indicator_ids = NULL,
                            prefix = NULL) {
  files <- sort(list.files(
    path       = image_dir,
    pattern    = pattern,
    full.names = TRUE,
    recursive  = FALSE
  ))

  if (!is.null(indicator_ids) && !is.null(prefix) && length(files) > 0) {
    file_ids <- sub(paste0("^", prefix, "(.*)\\.png$"), "\\1", basename(files))
    indicator_order <- indicator_ids[indicator_ids %in% file_ids]
    idx <- match(indicator_order, file_ids)
    files <- files[idx]
  }

  files
}

get_image_files2 <- function(image_dir,
                            pattern,
                            indicator_ids = NULL) {
  files <- sort(list.files(
    path       = image_dir,
    pattern    = pattern,
    full.names = TRUE,
    recursive  = FALSE
  ))

  if (!is.null(indicator_ids) && length(files) > 0) {
    # Strip dir + extension first: "Basic-CH_VACC_C_MSL-2018"
    base <- tools::file_path_sans_ext(basename(files))

    # Extract indicator between "Basic-" and "-YEAR"
    # → "CH_VACC_C_MSL"
    file_ids <- sub("^Basic-(.*)-[0-9]{4}$", "\\1", base)

    indicator_order <- indicator_ids[indicator_ids %in% file_ids]
    idx <- match(indicator_order, file_ids)

    files <- files[idx]
  }

  files
}

```

```{r}
image_dir <- normalizePath(
  file.path(source_path, middle_path, params$country, paste0(yr1, "-", yr2)),
  winslash = "/", mustWork = TRUE
)

# ad1 maps: needs special ordering by infolist$ID
files   <- get_image_files(
  image_dir   = image_dir,
  pattern     = "^ad1_map_.*\\.png$",
  indicator_ids = infolist$ID,
  prefix      = "ad1_map_"
)

# everything else: just sorted list.files

files_2 <- get_image_files(
  image_dir = image_dir,
  pattern   = "^ad2_map_.*\\.png$",
    indicator_ids = infolist$ID,
prefix      = "ad2_map_"
)

files_3 <- get_image_files(
  image_dir = image_dir,

  pattern   = "^overlay-.*\\.png$",
   indicator_ids = infolist$ID,
prefix      = "overlay-"
)

files_4 <- get_image_files(
  image_dir = image_dir,
  pattern   = "^scatter-.*\\.png$",
     indicator_ids = infolist$ID,
prefix      = "scatter-"
)

files_5 <- get_image_files(
  image_dir = image_dir,
  pattern   = "^ridge_.*\\.png$",
   indicator_ids = infolist$ID,
prefix      = "ridge_"
)

files_6 <- get_image_files(
  image_dir = image_dir,
  pattern   = "^interval-ad1-.*\\.png$",
   indicator_ids = infolist$ID,
prefix      = "interval-ad1-"
)

files_7 <- get_image_files(
  image_dir = image_dir,
  pattern   = "^interval-ad2-.*\\.png$",
    indicator_ids = infolist$ID,
prefix      = "interval-ad2-"
)

files_yr1 <- get_image_files2(
  image_dir     = image_dir,
  pattern       = paste0("^Basic-.*-", yr1, "\\.png$"),
  indicator_ids = infolist$ID
)

files_yr2 <- get_image_files2(
  image_dir     = image_dir,
  pattern       = paste0("^Basic-.*-", yr2, "\\.png$"),
  indicator_ids = infolist$ID
)
```


```{r, eval=FALSE}

image_dir <- normalizePath(file.path(source_path, middle_path, params$country),
                           winslash = "/", mustWork = TRUE)




files <- sort(list.files(path = image_dir,
                         pattern = "^ad1_map_.*\\.png$",
                         full.names = TRUE,
                         recursive = FALSE))
file_ids <- sub("^ad1_map_(.*)\\.png$", "\\1", basename(files))
indicator_order <- infolist$ID[infolist$ID %in% file_ids]
idx <- match(indicator_order, file_ids)
files <- files[idx]



files_2 <- sort(list.files(path = image_dir,
                         pattern = "^ad2_map_.*\\.png$",
                         full.names = TRUE,
                         recursive = FALSE))


files_3 <- sort(list.files(path = image_dir,
                         pattern = "^overlay-.*\\.png$",
                         full.names = TRUE,
                         recursive = FALSE))


files_4 <- sort(list.files(path = image_dir,
                         pattern = "^scatter-.*\\.png$",
                         full.names = TRUE,
                         recursive = FALSE))




files_5 <- sort(list.files(path = image_dir,
                         pattern = "^ridge_.*\\.png$",
                         full.names = TRUE,
                         recursive = FALSE))


files_6 <- sort(list.files(path = image_dir,
                         pattern = "^interval-ad1-.*\\.png$",
                         full.names = TRUE,
                         recursive = FALSE))


files_7 <- sort(list.files(path = image_dir,
                         pattern = "^interval-ad2-.*\\.png$",
                         full.names = TRUE,
                         recursive = FALSE))

files_yr1 <- sort(list.files(
  path       = image_dir,
  pattern    = paste0("^Basic-.*-", yr1, "\\.png$"),
  full.names = TRUE,
  recursive  = FALSE
))

files_yr2 <- sort(list.files(
  path       = image_dir,
  pattern    = paste0("^Basic-.*-", yr2, "\\.png$"),
  full.names = TRUE,
  recursive  = FALSE
))






```


## Overview

This report presents subnational prevalence estimates for `r length(indicatorlist)` indicators in `r country` using DHS data. A summary of all indicators and survey information is provided on the following page.

For each indicator, two sets of estimates are shown based on the latest and the previous DHS survey, when available. Details on the statistical methods used can be found at [https://sae4health.stat.uw.edu/method/approach_overview/](https://sae4health.stat.uw.edu/method/approach_overview/). The following outputs are included:

- **Sampling information**  
  Number of clusters, samples, and events by Admin-1 and Admin-2 areas.  

- **Admin-1 maps**  
  Weighted estimates at the Admin-1 level (without spatial smoothing). For each area, we present the point estimate, the width of the 90% uncertainty interval, and the probability that prevalence exceeds the national prevalence estimate for each survey.  

- **Admin-2 maps**  
  Spatially smoothed estimates at the Admin-2 level. Spatial smoothing is usually necessary to account for sparse data. The same set of maps as Admin-1 are presented.  

- **Admin-1 vs. Admin-2 comparison**  
  Side-by-side comparison of prevalence at Admin-1 and Admin-2. To improve readability, uncertainty intervals are omitted. Because Admin-2 estimates are spatially smoothed, the Admin-1 estimate may fall outside the range of the corresponding Admin-2 estimates, especially when prevalence is near 0% or 100%.  

- **Scatter plots (Admin-2)**  
  Comparison of weighted (unsmoothed) and smoothed Admin-2 estimates. Red dots indicate areas without valid weighted estimates. Smoothed estimates generally show shrinkage toward the mean and narrower uncertainty intervals. If the smoothed estimates appear have a much narrower range, as compared to the weighted estimates, this may indicate overshrinkage due to extreme data sparsity; such results should be interpreted with caution.  

- **Ridge plots (Admin-1)**  
  Visualization of the latest Admin-1 prevalence distributions and changes since the previous survey.  

- **Interval plots (Admin-1 and Admin-2)**  
  Prevalence estimates with uncertainty intervals for both the most recent and previous surveys. 


\newpage
## Country information

DHS surveys included `r c1 ` clusters in `r yr1` and `r c2 ` clusters in `r yr2`. Table 1 presents national estimates for 20 indicators from the two survey years. The subnational prevalence results are provided for `r ad11` Admin-1 areas and `r ad12` Admin-2 areas in this report.


```{r }
# The Rwanda 2025 DHS survey (Phase 9) is currently ongoing and its datasets will be available soon. 
# The Kenya 2025 interim DHS survey (Phase 9)
# The Nigeria 2023-24 DHS survey (Phase 8)
# The Ethiopia 2024–25 DHS survey (Phase 8)
# topic=surveyPrev::indicatorList[surveyPrev::indicatorList$ID %in% indicatorlist, c("ID","Topic")]
# 
# infolist=left_join(infolist,topic, by="ID" )
# 
# 
tab3=read.csv(file.path(source_path,"Gates-results/estimates", country, paste0(yr1, "-", yr2), "National3.csv"))
# 
# tab=read.csv(file.path(source_path,"Gates-results/estimates", country, "National.csv"))
# 
# tab <- tab %>%
#    dplyr::left_join(
#     infolist %>%  dplyr::select(Indicator = Description, Group, ID),
#     by = "ID"
#   ) %>%
#   dplyr:: arrange(Group, ID) %>%  # change Indicator to another column if needed
#  dplyr:: select(-ID)
# 
# pretty_names <- gsub("\\.", " ", names(tab)) 
# 
# # order groups like your sidebar and sort rows
# group_order <- c("Maternal Health", "Child Health",  "Social Determinants", "Health Outcomes")
# tab2 <- tab %>%
#   mutate(group = factor(Group, levels = group_order)) %>%
#   arrange(group)
# 
# # tab2$Indicator[4]="Percentage of children 12-23 months who had received MCV 1"
# # row counts per group for pack_rows()
# idx <- table(tab2$group)
# 
# 
# # pretty column names (remove dots)
# # pretty_names <- c("Indicator", "DHS 2013", "DHS 2018")
# 
# 
# pretty_names <- c("Indicator", paste0("DHS ", yr1), paste0("DHS ", yr2))
# 
# kable(tab2[1:3], #%>% dplyr::select(Indicator, DHS.2013, DHS.2018),
#       col.names = pretty_names,
#       align = c("l", "r", "r"),
#       booktabs = TRUE, caption = "National estimates.")%>%  # ) 
# kable_styling(latex_options = "hold_position",
#                 position = "center", 
#               
#               full_width = FALSE, 
#               stripe_color = "gray!10") %>%
#   row_spec(0, bold = TRUE) %>%
#   add_header_above(c(" " = 1, "Survey Year" = 2)) %>%
#   pack_rows(index = idx)   # <— adds “Child Health”, “Characteristics”, “Maternal Health” sections




```



```{r,eval=T , echo=FALSE, results='asis'}

tab=read.csv(file.path(source_path,"Gates-results/estimates", country, paste0(yr1, "-", yr2), "National.csv"))
tab[5,3]<-tab[5,3]*10
tab[5,2]<-tab[5,2]*10
tab <- tab %>%
   dplyr::left_join(
    infolist %>%  dplyr::select(Indicator = Description, Group, ID),
    by = "ID"
  )
# force row order by infolist$ID (and keep your group order)
group_order <- c("Maternal Health", "Child Health", "Social Determinants", "Health Outcomes")

tab2 <- tab %>%
  mutate(
    Group = factor(Group, levels = group_order),
    ID    = factor(ID, levels = infolist$ID)     # <-- key: order by infolist$ID
  ) %>%
  arrange(Group, ID)                             # group order first, then ID order
idx <- table(tab2$Group)  

pretty_names <- c("Indicator", paste0("DHS ", yr1), paste0("DHS ", yr2))
# 
#   knitr::kable(
#   tab2 %>% dplyr::select(Indicator, DHS.2010, DHS.2021),   # <-- change cols to yours
#   col.names = pretty_names,
#     format="latex",
#   align = c("l", "r", "r"),
#   booktabs = TRUE,
#   caption = "National estimates (%) "
# ) %>%column_spec(1, width = "10cm") %>% kable_styling(
#              latex_options = "hold_position",
#                 # latex_table_env = "tabularx",
#                 position = "left",
#                 full_width = FALSE,
#                 stripe_color = "gray!10") %>%
#   row_spec(0, bold = TRUE) %>%
#   add_header_above(c(" " = 1, "Survey Year" = 2)) %>%
#   pack_rows(index = idx)


# kable(tab2[1:3], #%>% dplyr::select(Indicator, DHS.2013, DHS.2018),
#       col.names = pretty_names,
#       align = c("l", "r", "r"),
#       booktabs = TRUE, caption = "National estimates.")%>%  # ) 
# kable_styling(latex_options = "hold_position",
#                 position = "center", 
#               
#               full_width = FALSE, 
#               stripe_color = "gray!10") %>%
#   row_spec(0, bold = TRUE) %>%
#   add_header_above(c(" " = 1, "Survey Year" = 2)) %>%
#   pack_rows(index = idx)   # <— adds “Child Health”, “Characteristics”, “Maternal Health” sections

tab2 %>%dplyr::select(Indicator, dplyr::starts_with("DHS.")) %>%
  knitr::kable(
    format = "latex",
    col.names = pretty_names,
    align = c("l", "r", "r"),
    booktabs = TRUE,
    caption = "National prevalence (values in \\%; child mortality indicators are reported per 1,000)."
  ) %>%column_spec(1, width = "13cm") %>%
  kableExtra::kable_styling(
    latex_options =  c("hold_position", "caption_below"),
    position = "left",
    full_width = FALSE
  ) %>%
  kableExtra::row_spec(0, bold = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "Survey Year" = 2)) %>%
  kableExtra::pack_rows(index = idx)


```


```{r,eval=FALSE}

tab <- read.csv(file.path(source_path,"Gates-results/estimates", country, "National.csv")) %>%
  left_join(infolist %>%  dplyr::select(Indicator = Description, Group, ID), by = "ID")

group_order <- c("Maternal Health", "Child Health", "Social Determinants", "Health Outcomes")

tab2 <- tab %>%
  mutate(
    Group = factor(Group, levels = group_order),
    ID    = factor(ID, levels = infolist$ID)
  ) %>%
  arrange(Group, ID)

idx <- table(tab2$Group)

pretty_names <- c("Indicator", paste0("DHS ", yr1), paste0("DHS ", yr2))

kbl(
  tab2 %>%  dplyr::select(Indicator, DHS.2010, DHS.2021),
  format = "latex",            # <-- add
  booktabs = TRUE,
  col.names = pretty_names,
  align = c("l", "r", "r"),
  caption = "National estimates (%) "
) %>%
  add_header_above(c(" " = 1, "Survey Year" = 2)) %>%
  pack_rows(index = idx) %>%
  kable_styling(
    latex_options   = "hold_position",
    latex_table_env = "tabularx",   # <-- key fix (avoids tabu)
    position = "center",
    full_width = TRUE
  ) %>%
  column_spec(1, width = "15cm") %>%  # keep this
  row_spec(0, bold = TRUE)

```



```{r, out.width='0.8\\linewidth', fig.align='center',fig.cap="Admin-1 and Admin-2 map ", eval=FALSE}



ad1=ggplot(poly.adm1) +
  geom_sf(fill = NA, color = "black", size = 0.3) +
 theme_bw() +
  theme(
    axis.text = element_blank(),      # remove degree labels
    axis.ticks = element_blank(),     # remove tick marks
    axis.title = element_blank(),     # remove axis titles
    panel.grid = element_blank(),     # remove grid lines
    panel.border = element_rect(color = "black", fill = NA)  # keep outer frame
  )
ad2=ggplot(poly.adm2) +
  geom_sf(fill = NA, color = "black", size = 0.3) +
  theme_bw() +
  theme(
    axis.text = element_blank(),      # remove degree labels
    axis.ticks = element_blank(),     # remove tick marks
    axis.title = element_blank(),     # remove axis titles
    panel.grid = element_blank(),     # remove grid lines
    panel.border = element_rect(color = "black", fill = NA)  # keep outer frame
  )

ad1+ad2
```





```{r, echo=FALSE, results='asis' , out.width='\\ linewidth'}
latex_img <- function(p, width="\\linewidth") {
  if (length(p) == 0 || is.na(p) || !nzchar(p) || !file.exists(p)) {
    cat(sprintf("\\textit{(missing file) %s}\\\\\n\n", ifelse(length(p), p, "")))
    return(invisible())
  }
  cat(sprintf("\\begin{center}\\includegraphics[width=%s]{%s}\\end{center}\n\n",
              width, p))
}



indicatorlists=indicatorlist
for (i in seq_along(indicatorlists)) {
  cat("\\newpage\n\n")

id  <- indicatorlists[i]
def <- infolist$Full_definition[infolist$ID == id]
short <- infolist$Description[infolist$ID == id]

cat(sprintf("## %s\n\n", short))

cat(sprintf("`%s`: %s\n\n", id, def))

  
cat("\\textbf{Sampling information for previous survey}\n\n")
    cat(infolist[infolist$ID == indicatorlists[i], ]$Description, "\n\n", sep = "")
      cat(paste0(tab3[tab3$ID==indicatorlists[i], ]$previous_no_ad2," Admin-2 areas have no samples, ",
                 tab3[tab3$ID==indicatorlists[i], ]$previous_fixed)," Admin-2 areas have unstable variance." , "\n\n", sep = "")
    
  if (i <= length(files_yr1)) latex_img(files_yr1[i],  ".75\\linewidth")
  cat("\\newpage\n\n")

  cat("\\textbf{Sampling information for current survey}\n\n")
   cat( infolist[infolist$ID == indicatorlists[i], ]$Description, "\n\n", sep = "")
         cat(paste0(tab3[tab3$ID==indicatorlists[i], ]$current_no_ad2," Admin-2 areas have no samples, ",
                 tab3[tab3$ID==indicatorlists[i], ]$current_fixed)," Admin-2 areas have unstable variance." , "\n\n", sep = "")
    
   
  if (i <= length(files_yr2)) latex_img(files_yr2[i],  ".75\\linewidth")
  cat("\\newpage\n\n")

  cat("\\textbf{Maps for Admin-1 unsmoothed weighted estimates}\n\n")
   cat(infolist[infolist$ID == indicatorlists[i], ]$Description, "\n\n", sep = "")
  if (i <= length(files))     latex_img(files[i],  ".8\\linewidth")
  cat("\\newpage\n\n")

  cat("\\textbf{Maps for Admin-2 smoothed weighted estimates}\n\n")
   cat(infolist[infolist$ID == indicatorlists[i], ]$Description, "\n\n", sep = "")

  if (i <= length(files_2))   latex_img(files_2[i], ".8\\linewidth")
   
     cat("\\newpage\n\n")
       cat("\\textbf{Overlay plots for Admin-1 unsmoothed and Admin-2 smoothed weighted estimates}\n\n")
  if (i <= length(files_3))   latex_img(files_3[i], ".8\\linewidth")
       cat("\\textbf{Scatter plots for Admin-2 unsmoothed and smoothed weighted estimates }\n\n")

  if (i <= length(files_4))   latex_img(files_4[i], ".7\\linewidth")
   
       cat("\\newpage\n\n")
  cat("\\textbf{Ridge plots for Admin-1 results}\n\n")
  cat(infolist[infolist$ID == indicatorlists[i], ]$Description, "\n\n", sep = "")
  cat(paste0( "Left: ", yr2, " weighted estimates, Right: Changes from ",yr1), "\n\n", sep = "")
  if (i <= length(files_5))   latex_img(files_5[i], ".9\\linewidth")

  
       cat("\\newpage\n\n")

    cat("\\textbf{Interval plots for Admin-1 weighted estimates}\n\n")
  cat(infolist[infolist$ID == indicatorlists[i], ]$Description, "\n\n", sep = "")
  if (i <= length(files_6))   latex_img(files_6[i], ".9\\linewidth")
    
   cat("\\newpage\n\n")

     cat("\\textbf{Interval plots for Admin-2 smoothed weighted estimates}\n\n")
  cat(infolist[infolist$ID == indicatorlists[i], ]$Description, "\n\n", sep = "")
  if (i <= length(files_7))   latex_img(files_7[i], "\\linewidth")

    
   
}
```




